% Chapter 1

\chapter{Dark Energy and Modified Gravity \label{chap:DE-MG-Overview}} % Main chapter title

 % For referencing the chapter elsewhere, use \ref{Chapter1} 

%----------------------------------------------------------------------------------------

% Define some commands to keep the formatting separated from the content 
\newcommand{\keyword}[1]{\textbf{#1}}
\newcommand{\tabhead}[1]{\textbf{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\file}[1]{\texttt{\bfseries#1}}
\newcommand{\option}[1]{\texttt{\itshape#1}}

%----------------------------------------------------------------------------------------
%

The theoretical framework to describe gravity and space-time in cosmology is Einstein's General Relativity which is
one of the cornerstones of modern physics.
Einstein formulated this theory more than 100 years ago, without the purpose
of explaining cosmology, but mostly to solve the theoretical challenges posed by his relativistic mechanics under the influence of a gravitational field. Besides maybe the
perihelion of Mercury there was no experimental need for it. It was more than a decade later, thanks to observations
by Hubble and Slipher, that physicists where convinced that there were objects much farther away from
our galaxy and that these objects were receding away from us.

%At this point the field of physical cosmology was born and the theoretical basis for it,
%based on General Relativity, was developed by Lema\^{\i}tre, Friedmann, and Einstein himself among other
%notable scientists.

One century later, General Relativity has passed numerous very stringent tests; from laboratory experiments (\cite{cite}) , to
low orbit tests (\cite{cite}) and solar system tests (\cite{cite}), to pulsar timing tests and the recent
exciting first detection of gravitational waves (\cite{cite}).
It is impressive that a theory that was formulated on the grounds of some very basic
principles, has proven to be so accurate across several orders of magnitude in scales.

To explain the accelerated expansion of the Universe, which was observationally verified almost 20 years ago,
(\cite{cite, supernova, 1998}), Einstein's General Relativity needs to invoke a cosmological constant, which despite the fact that
it is allowed by the theory (see Lovelock's theorem below in\cref{sub:Einstein-Hilbert} ),
possesses many unsatisfactory properties. 
We will review this issue in \cref{sub:CC-problem} below.

In order to provide some context for the main results of this dissertation, we will
provide in the following sections a very brief overview of the standard cosmological model. 
In \cref{sec:GR-framework} we will review the main principles and the mathematical 
formulation of General Relativity, its field equations and
its linearized Newtonian limit.
In \cref{sec:Standard-LCDM} we will deal with the composition of the Universe,
its evolution and the standard cosmological scenario.


\section{The framework of General Relativity \label{sec:GR-framework}}

%For Einstein, Special Relativity (SR) was unsatisfactory and didn't represent a complete theory, because it dealt with 
%intertial frames and under the influence of
%a gravitational field, objects would be accelerated. Moreover due to the equivalence of mass and energy in SR, 
%an object with a high kinetic energy in horizontal direction would have to be accelerated differently towards the Earth, than an object
%with a smaller velocity, therefore violating Newtonian observations, which claim that all test bodies experience
%the same acceleration in a gravitational field, regardless of its velocity or composition.
%This observation that the inertial and gravitational masses must be equivalent was of striking significance for Einstein, 
%which elevated it to a guiding principle in the construction of a relativistic theory of gravity (see \cite{pais, wald, bartelmann}).

%The really ingenious step was connecting this principle with differential geometry. Then, instead of the rigid 
%Euclidean picture, space and time would become a dynamical 4-dimensional structure, described by differential manifolds.
%Gravity would curve this space-time, but in order to maintain the equivalence principle, it is always possible to transform away gravity 
%and map it to a flat Euclidean space, which is precisely one of the properties of a differentiable manifold.
%If one studies the consequences of this simple principle, it leads, without the need of formulating a theory, 
%to concepts like gravitational redshift and gravitational light deflection. 
%Constructing from there a fully-fledged non-linear theory of gravity took Einstein more than 10 years, with the (maybe indirect) help
%of very bright mathematicians like Grossmann, Levi-Civita and Hilbert.

Although Einstein based its construction of General Relativity on considerations of the \emph{Equivalence Principle},  we can say, in more modern terms, that General Relativity is a theory of 
a dynamical tensor field, the metric $g_{\mu \nu}$, which defines the lengths 
of space-time intervals $ds^2 = g_{\mu \nu} dx^{\mu} dx^{\nu}$ and which is invariant under diffeomorphisms. 
All particles and fields couple to the metric $g$ in a universal way.
This means that the equations of motion and all the physical properties do not depend on the chosen coordinates.
Diffeomorphism invariance is a very important symmetry that has to be respected if one wishes to construct
extensions of gravity. 
%However, at the smallest (Planck length) and largest (super-horizon) scales, there are suggestions
%that this symmetry might be broken in order to be able to construct a consistent theory of Quantum Gravity
%(\cite{cite Rovelli}).



\subsection{The Einstein-Hilbert action and the field equations \label{sub:Einstein-Hilbert}}

Although Einstein postulated the field equations of General Relativity in a heuristic
form, they can be obtained by varying the so-called Einstein-Hilbert action
\begin{equation}\label{eq:Einstein-Hilbert action}
S = \frac{1}{16 \pi G} \int \textrm{d}x^4 \sqrt{-g}\left( R - 2\Lambda + \mathcal{L}_m \right) \quad 
\end{equation}
with respect to the metric $g_{\mu \nu}$.
Here, $R=R^\mu_\nu$ is the Ricci scalar and $R^\munu$ is the Riemann tensor (see standard GR textbooks like \cite{wald} for its definition), 
which are functions of derivatives of the metric. The volume element is defined as $\textrm{d}x^4 \sqrt{-g}$, where
$\sqrt{-g}$ is the square root of the determinant of the metric.
Furthermore, the cosmological constant is $\Lambda$, the gravitational constant is $G$ and $\mathcal{L}_m $ is the Lagrangian of matter and radiation species.
Then the variation ($\delta S / \delta g_{\mu \nu} = 0$), yields the field equations:
\begin{equation}\label{eq:Einstein-field-equations}
G_\munu + g_{\mu \nu} \Lambda = 8 \pi G T_{\mu \nu} \quad ,
\end{equation}
where the Einstein tensor is defined as $G_\munu \equiv R_{\mu \nu} - \frac{1}{2}g_{\mu \nu} R \;$ and the 
energy-momentum tensor is defined as:
\beeqp$
T_\munu = -\frac{2}{\sqrt{-g}} \frac{\delta \mathcal{L}_m}{\delta g_\munu}
$
What \cref{eq:Einstein-field-equations} expresses is that geometry (and therefore the dynamics of the metric) is sourced by 
the energy and momentum of the fields living on this manifold.
As it was famously expressed by \cite{Misner, Wheeler, Gravitation}:
"Matter tells space-time how to curve and space-time tells geometry how to move" .

Another important property of these field equations, is that due to a purely geometric property called the Bianchi identity, which
states that the covariant divergence of the Einstein tensor is identically zero:
$\nabla_{\mu} G^\munu = 0$, 
we can ensure that the energy-momentum tensor is locally conserved:
\beeqc$
\nabla_{\mu} T^\munu = 0
$
therefore one recovers all the well-known properties of classical and fluid mechanics.

%Einstein's field equations are very complicated to solve analytically in its full glory. They consist on
%10 coupled non-linear partial differential equations.
%There are of course many interesting properties, solutions and implications of these equations, 
%which are out of the scope of this work.
%We refer the interested reader to very good textbooks like \cite{Carroll, Schultz, Misner, Thorne, Wheeler, Wald}.
%The richness and complexity of this theory is one of the reasons it has to be so successful explaining 
%star formation, gravitational collapse, black holes, gravitational waves and the evolution of the Universe.

%\subsubsection{Lovelock's theorem and the uniqueness of General Relativity}

One might then ask if these field equations are unique, especially if as in our case, 
we are interested in testing these equations at the very largest scales of the Universe and we might 
be interested in modifying General Relativity in order to match current observations.
Thanks to a theorem by Vermeil and Cartan (\cite{cite Vermeil, Cartan, 1921}) 
and further simplified by Lovelock (\cite{1970, Lovelock}) we can state that in 4 dimensions,
the only divergence-free, rank-2 tensor $\mathcal{G}$, which depends on at most second derivatives
of the metric must be of the form:
\beeqp$
\mathcal{G} =  \alpha R_{\mu \nu} + \left( \Lambda - \frac{\alpha}{2} R \right) g_{\mu \nu}
$
Therefore, to ensure the correct Newtonian limit of the theory (Newtonian Poisson equation), we must set $\alpha \equiv 1$
and the proportionality constant between the Einstein tensor $G_\munu$  and the energy-momentum tensor
$T_\munu$ has to be set to $8 \pi G$.
This theorem has fundamental implications for cosmology, which we will comment 
in the following sections, when needed.
%First of all, it states that the Cosmological Constant (CC) has to appear in the classical theory of gravity,
%therefore, if ---as we will see below in \cref{sub:CC-problem})--- we are not satisfied with the CC as an explanation
%of the accelerated expansion of the Universe, we still have to explain why $\Lambda$ disappears from the Einstein's field equations.
%Second, this theorem gives a sort of roadmap to look for modifications of gravity, which can account for the late-time expansion of the Universe.
%One option is to go beyond 4 dimensions (as in DGP models \cite{DGP models}),
%violate the conservation of the energy-momentum tensor with some exotic species or
%add extra degrees of freedom, since with the metric alone we can't construct a more general theory.
%This last option has led to the recent interest in modified gravity theories with a scalar field (see \cite{cite some scalar field}), 
%with vector fields (see \cite{cite}) and with the addition of several metric (tensor) fields (see \cite{cite}). 


%----------------------------------------------------------------------------------------

\section{The standard cosmological model \label{sec:Standard-LCDM}}

The cosmological principle states
that no observer in the Universe is special and that each observer sees the Universe in the same way independently of
spatial rotations, therefore space-time has to be described by a homogeneous and isotropic metric $g_\munu$.
If furthermore, we can define a foliation of space-time, in which there is a preferred timelike direction, orthogonal
to the spatial hypersurfaces, we end up with a metric which solves the Einstein's field equations \ref{eq:Einstein-field-equations};
the so-called Friedmann-Lema\^{\i}tre-Robertson-Walker (FLRW) metric:
\beeqc$\label{eq:FLRW-full}
ds^2 = g_\munu dx^\mu dx^\nu = -dt^2 + a^2 (t) d\varsigma^2 
$
where $a$ is the scale factor, $t$ the cosmic time coordinate, and $d\varsigma^2$ is the time-independent spatial metric:
\beeqc$\label{eq:FLRW-spatial}
d\varsigma^2  = \gamma_{i j} dx^i dx^j = \frac{dr^2}{1-kr^2} + r^2(d\theta^2 + \sin^2 \theta d \phi^2)
$ 
where $r$ is the radial coordinate, $\theta$ the polar angle and $\phi$ the azimuthal angle.
The curvature $k$ which can be 0, 1 or -1, corresponds to Universes which are flat, closed or open, respectively.
Due to the stringent constraints on $k$ given by recent observations,
we will use for the rest of this work only a flat geometry with $k=0$.
Also it is standard convention that "Greek" indices run from 0 to 3, as in \cref{eq:FLRW-full},
while "Latin" indices run from 1 to 3 as in equations involving only spatial coordinates, like \cref{eq:FLRW-spatial}.


\subsection{The Friedmann equations \label{sub:Friedmann-eqs}}

The FLRW metric \cref{eq:FLRW-full} due to its scale factor, which is dependent on time, implies immediately that the Universe can expand
in its spatial coordinates. The equations describing the evolution
of the scale factor are called the \emph{Friedmann} equations. To derive them,
we need to introduce in the right hand side of Einstein's equations \cref{eq:Einstein-field-equations} an energy-momentum tensor of a perfect fluid:
\beeqc$
T^\munu = (\rho + p)u^\mu u^\nu + p g^\munu
$
which is justified since at cosmological scales, we expect the background matter in the Universe to be absent of dissipative and viscous forces.
The density $\rho$ and the pressure $p$ are the sum of the densities and pressure terms of all matter and radiation species in the universe.
If we write down now the (00) and ($i i$) components of the Einstein's field equations \cref{eq:Einstein-field-equations}, computing
the Riemann and Ricci tensors of a FLRW metric, 
we end up with two ordinary differential equations for the scale factor $a(t)$:
\beeqal$
\left(\frac{\dot a}{a} \right)^ 2 &= \frac{8 \pi G}{3} \rho + \frac{\Lambda}{3}  \label{eq:Friedmann-1st}\\
\frac{\ddot a }{a} &= -\frac{4 \pi G}{3} (\rho +3 p) + \frac{\Lambda}{3} \label{eq:Friedmann-2nd}
$
These are the so-called Friedmann equations, which determine the evolution of the Hubble function defined as:
\beeqp$
H(t) \equiv \frac{\dot a(t)}{a(t)}
$
The critical density of the Universe is defined as:
\beeqc$
\rho_{cr} = \frac{3 H^2}{8 \pi G}
$
which is the critical density that an Universe with zero curvature $k=0$ would have according to the observed value of the Hubble function.
So, for each species in the Universe, with energy density $\rho_i$, we can define the energy density fraction as:
\beeqc$
\Omega_i (t) = \frac{\rho_i (t)}{\rho_{cr} (t)}
$ 
so that the first Friedmann equation \cref{eq:Friedmann-1st} (for a flat Universe) can be written as:
\beeqp$
\sum_i \Omega_i (t) = 1
$
Another important quantity for each matter species is its equation of state:
\beeqc$
w \equiv \frac{p}{\rho}
$
which enters in its cosmological evolution equation.


\subsection{The $\mathbf{\Lambda}$CDM model \label{sub:LCDM}}

As we have seen before, the field of cosmology is relatively new, with less than 100 years of theoretical development
and even much less time of observational progress. 
However, in the last few decades, with the precise measurements of the Cosmic Microwave Background (CMB) radiation,
the rapid progress in galaxy surveys and the impressive development of cosmological simulations, an observationally very successful standard model has emerged, the so-called
$\lcdm$ model.

In this model, and according to the latest observations (see \cite{planck_collaboration_planck_2016-1}),
almost 70\% of the energy density of the Universe is composed by vacuum energy, attributed to the Cosmological Constant
$\Lambda$, 25\% by Cold Dark Matter and less than 5\% by baryons. The remaining components are photons ---which are basically negligible today, despite the amount of light and radiation in the Universe--- and massive neutrinos,
whose mass and therefore its contribution to the "cosmic pie" has not been measured precisely enough yet.
%Up to 1965?, with the discovery of the Cosmic Microwave Background (CMB) radiation by \cite{Penzias and Wilson},
%cosmology was a mere speculative science. 
%Only with the confirmation of the "Hot Big Bang paradigm" by the measurement of the radiation left over
%380,000 years after the Big Bang, is that physicists started to make precise descriptions about the composition and 
%the evolution of the matter species composing the Universe. The first CMB observations clarified that the Universe started
%as a hot plasma of electrons, protons and photons, whose interactions in an expanding space-time matched very well the observed properties
%of the present Universe.

%However, there was always a piece missing in the puzzle, since the observed densities of visible matter in the Universe were too low.
%For many years, there were big discussions about open and closed Universes and about invisible forms of matter.
%Only with the amazing development of galaxy surveys in the 1980's and precise measurements of stars velocities in the Milky Way and other galaxies
%and the development of cosmological N-body simulations in the 1990's, most cosmologists were convinced that there had to be a form of matter in the Universe
%that does not interact with baryons or with electromagnetic fields, but it interacts in a standard way with gravitational fields (\cite{cite;
%recent peebles, Dolag DM review, reviews, books}).
%They called it "Cold Dark Matter" (CDM) because of its apparent vanishing pressure and its preference for clustering into big structures.
%
%However, there were still some inconsistencies between the growth of structures in a pure 'CDM+baryons' model, predicted
%by simulations and the observed structures in the Universe (see \cite{reviews Dolag, peacock}), so that many scientists were claiming for the addition
%of a cosmological constant. However, this claims were not significant until in 1998 the teams led by \cite{Perlmutter and ..} used Supernova Type Ia
%observations to prove that the Universe was experiencing a phase of accelerated expansion. This led to the re-introduction of the Cosmological Constant
%and due to the theoretical problems posed by its so small observed value, to the development of a whole new 
%field of research: Modified Gravity and Dark Energy.

%This very rough and quick overview explains the name of the current standard cosmological model: the $\lcdm$ model. 
This special mixture of cosmological ingredients in the Universe, leads to a well defined background evolution of the Hubble function $H(z)$. However, as we will see below, the Cosmological Constant is not very satisfactory from the theoretical point of view.
Therefore if we leave open the possibility that the accelerated expansion of the Universe is caused by a "Dark Energy" component, with 
an unknown equation of state $w_{DE}(z)$ (where $w_{DE}(z) = -1$ would correspond to the Cosmological Constant), then together with the "standard" matter species (with $w=0$ for CDM and baryons and $w=1/3$ for radiation), we can write down
the evolution of the Hubble parameter as a function of redshift $z \equiv 1/a - 1 $ :
\beeqalsp$\label{eq:H-of-z-lcdm}
H^2 (z) =& H_0 \left(  \Omega_c (1+z)^3 + \Omega_b (1+z)^3  + \Omega_r (1+z)^4   
\phantom{\exp \left[ \int_{0}^{z}  \frac{3}{1} \right]} \right. \\
      & \left.  +\Omega_{DE}  \exp \left[ \int_{0}^{z} \dx{}{\tilde z} \frac{3(1+w_{DE}(\tilde z))}{1+\tilde z} \right]     \right)
$
where we have left out neutrinos, since their evolution at high and low redshifts is not so straightforward to write, due to their changing from relativistic to non-relativistic particles in the history of the Universe.
If we assume a constant $w$ and a single matter species, one can analytically solve 
\crefrange{eq:Friedmann-1st}{eq:Friedmann-2nd} and find that the scale factor behaves 
as:
\beeqp$
a \propto t^{\frac{2}{3(1+w)}}
$
Since each species evolves with a different power of the scale factor, there 
were three different epochs in the evolution of the Universe, the radiation dominated era (RDE) in which $a \propto t^{1/2}$, the matter domination era
(MDE) in which $a \propto t^{2/3}$ (also called the Einstein-de Sitter Universe) and finally the future dark energy dominated era, also called the de Sitter regime, in which
the total energy density is constant and therefore $a \propto \exp(H t)$.
From the Hubble function, one can obtain measurable 
cosmological distances to astrophysical objects and the
the age and the size of the observable Universe.

Since the Universe is expanding and we measure astrophysical objects using mainly electromagnetic radiation (except after 2016 \cite{cite LIGO}, 
where Gravitational Wave Astronomy was born), it is important to define certain distances that can be measured in cosmological observations.
Light follows null-geodesics, so that under an FLRW metric as in \cref{eq:FLRW-full}, 
light satisfies the equation:
\beeqc$
- c dt^2 + a^2 (t) d\varsigma^2 = 0 
$
where we have recovered the speed of light $c$ to avoid confusion.
Solving for $\varsigma$ and therefore integrating this equation, defines the comoving distance $d_c$:
\beeqp$
d_c \equiv \int_0^{\varsigma_1} \mathrm{d}\varsigma = - \int_{t_0}^{t^1} \frac{c}{a(t)} \dx{}{t} 
$
Since $H = (\mathrm{d}a/\mathrm{d}t)/a = (\mathrm{d}(1/(1+z))/\mathrm{d}t)/(1/(1+z)) $, then $\mathrm{d} t = -\mathrm{d}z / ((1+z) H) $,
so that the comoving distance can be defined as:
\beeqp$
d_c (z) = \int_{0}^{z} \frac{\dx{}{\tilde z}}{H(\tilde z)}
$
Making similar geometrical considerations (which we will not detail here, see \cite{amendola_dark_2010}), one can find the
luminosity distance:
\beeqc$
d_L = (1+z) d_c
$
and the angular diameter distance:
\beeqc$
d_A = \frac{d_c}{1+z}
$
where we have set the curvature $k$ a priori to zero.


\subsection{The cosmological constant problem \label{sub:CC-problem}}

%There are two main problems with the $\lcdm$ model: The first one is $\Lambda$; the second
%one, CDM.
%We will not deal here with the CDM problem, but it is just worth to say, 
%that despite large efforts to find 
%the CDM particle, nothing significant has been found so far.
After more than two decades of intensive experimental and observational searches,
physicists are still not clear about what constitutes almost 95\% of the energy budget of
the Universe. These are the so-called Dark Matter and Dark Energy problems.

\subsubsection*{The fine-tuning problem}
While a Cosmological Constant can explain very well observations so far, the Cosmological Constant (CC) problem is a more profound one, since it 
involves a missing understanding of both Quantum Field Theory and General Relativity, as we will see now in more detail.
From \cref{eq:Friedmann-1st}, we see that under dark energy domination, 
the CC is of the order of the square of the Hubble parameter today:
\beeqp$
\Lambda \approx H_0^2 = (2.1 h \times 10^{-42} \mathrm{GeV})^2
$
so that as an energy density $\rho_{\Lambda} = \Lambda / 8\pi G$ we would obtain:
\beeqc$\label{eq:rho-Lambda}
\rho_{\Lambda} \approx 10^{-47} \mathrm{GeV}^4
$
where we have used $1/G = m_{Pl} $ and the Planck mass is equal to $m_{Pl} = 10^{19} \mathrm{GeV}$.
Making a very rough quantum field theory calculation, one can see that if the vacuum energy density comes from the zero point energy of a single field with mass $ m $ and momentum $ k $, with energy
$E = \sqrt{k^2 + m^2}/2$,
then one can sum all contributions from all momenta up to a cut-off scale $k_{\rm max}$:
\beeqc
 $\langle \rho_{\rm vac} \rangle = \int_{0}^{k_{\rm max}} \frac{4\pi k^2 \dx{}{k}}{(2\pi)^3}\frac{\sqrt{k^2 + m^2}}{2} \approx \frac{k_{\rm max}^4}{16 \pi ^2} 
$
where we have used the fact that the integral will be dominated by large modes ($k \gg m$).
If we take this cut-off scale to be of the order of the Planck mass $m_{Pl}$, which is a scale up to which we believe GR might be still valid, then we find
\beeqc$\label{eq:rho-vac-rough}
\langle \rho_{\rm vac} \rangle \simeq 10^{74} \mathrm{GeV}^4
$
which is $10^{121}$ larger than the value found above in \cref{eq:rho-Lambda}.
This is the famous ``fine-tuning`` problem: If we want to reconcile the measured 
value of $\Lambda$ with the expected vacuum energy provided by quantum fields, 
there has to be a cancellation which is exact to 120 orders of magnitude.

However, we must stress here again that the result in \cref{eq:rho-vac-rough} 
is a very rough calculation
that does not consider the symmetries of the problem and does not respect the equation of state
of vacuum energy: doing the above rough calculation also for the pressure 
would yield $\langle p_{\rm vac} \rangle / \langle \rho_{\rm vac} \rangle  = 1/3$, instead of the expected $w=-1$ (see \mcite{Jerome, Martin}).
Therefore, under a proper regularization scheme, the vacuum energy density is calculated to be 
$\langle \rho_{\rm vac} \rangle \approxeq 10^{10} \mathrm{GeV}^4$, which is more or less "just"
50 orders of magnitude larger than $\rho_{\Lambda}$ (see \mcite{Jerome, Martin}).
Though this cancellation could be realized by different mechanisms, the problem is worse than that;
the fine-tuning has to be performed at each subsequent order in quantum field perturbation theory (see the comprehensive review by \mcite{Jerome, Martin}). 
One has to find a powerful symmetry that
"screens" the vacuum energy from gravitating and affecting the Einstein's field equations.

\subsubsection*{The coincidence problem}

Apart from the unsatisfactory discrepancy between the Cosmological Constant and the expected 
vacuum energy density, the value of $\Omega_{\Lambda}$ today is suspiciously close to the 
value of $\Omega_m$, for no apparent reason, and this has been the case only very recently in cosmological time scales. The redshift at which both energy densities coincide ($z_{co}$) is:
\beeqp$
z_{co} = \left( \frac{\Omega_{\Lambda}}{1-\Omega_{\Lambda}}  \right)^{\frac{1}{3}} -1$
(see \mcite{Luca, book}).
Therefore, for a value today of $\Omega_{\Lambda}=0.7$, the coincidence redshift is $z_{co}\approx 0.3$.
This is indeed a very recent time and depends strongly on the ratio of $\Omega_{\Lambda}/\Omega_{m}$:
if this ratio was just 10 times smaller or larger, we as observers would not measure accelerated expansion today.

Several models have been proposed so far to deal with this apparent coincidence, for example
Dark Energy models with tracker and scaling solutions (see \mcite{Luca, book}) in which $\rho_{DE}$ catches the trend of $\rho_{m}$ no matter which initial conditions have been chosen
or where one constructs an equation of state for DE that changes just very recently in time.
Other models try to link the onset of acceleration by connecting it with the evolution of non-linear
structure formation, for example with backreaction effects (\mcite{Buchert}), 
but these models have not been so successful so far.

In \cref{sub:GNQ} we will see how a coupling of the scalar field to the mass of neutrinos 
could link the onset of acceleration to the neutrinos becoming non-relativistic and therefore
it can alleviate the coincidence problem.
 
Other scientists resort to the anthropic principle (\mcite{Luca, book, Weinberg})
and claim that for us observers to exist in a Universe with galaxies, stars and planets
and to have evolved sufficiently to develop advanced civilizations, the Cosmological Constant
can only have certain precise range of values \mcite{Weinberg}. While it is true that these constraints are applicable
to our case, this explanation is not widely accepted as a "physical" solution of the problem.


\subsection{\revtext{Gravitational potentials} (or Linearized Einstein Equations)}

Since we observe non-homogeneities in the Universe around us, we need to go beyond
the background description of General Relativity given by the Friedmann equation \ref{eq:H-of-z-Lcdm}.
Observations of the CMB tell us that the initial fluctuations are small ---of the order of $10^{-5}$---
and therefore we can split the metric $g_\munu$ in a background part $\bar{g}_\munu$ 
and a perturbation part $\delta{g}_\munu$.

Introducing the conformal time $\mathrm{d}\tau = \dx{}{t}/a $ and using the FLRW metric \cref{eq:FLRW-full} as the
background one, the perturbed metric can be written as:
\beeqalsp$
ds^{2} &= a^2(\tau) \left(-(1+2\Psi)d\tau^{2} + (B_{,i} + S_i)d\tau dx^i \right. \\
       & \left. -  (-2\Phi \delta_{ij} + 2E_{,ij} + F_{i,j} + F_{j,i} + h_{ij}) \right)
$
where we have decomposed the metric into 4 scalar functions $\Phi$, $\Psi$,
$E$ and $B$, two vectors $F_i$ and $S_i$  and a tensor $h_{ij}$ perturbation.
Moreover the vector perturbations satisfy the constraints $F^i_{,i} =0$ and 
$S^i_{,i} =0$ (divergence-free) and the tensor perturbation satisfies the traceless
and transverse constraints $h^i_i = 0$, $h^i_{j,i} = 0$.
Therefore, counting the total number of degrees of freedom, we end up with 10 independent
components, which are the same independent components that a (symmetric) metric $g_\munu$ has.

Vector perturbations decay very quickly and therefore are uninteresting in standard cosmology. Tensor
perturbations describe gravitational waves, and though there is an increased interest in them due
to gravitational wave detection \mcite{LIGO} and B-mode polarization in the CMB \mcite{Planck, BICEP},
we will not treat them in this work.

Moreover, it can be shown that under a change of coordinates $x^\mu \rightarrow x^\mu + \xi^\mu$, there is still a gauge freedom that can be fixed and therefore only 6 degrees of freedom are really independent. In the following we will choose the \emph{conformal Newtonian gauge }, which is defined
by $E_i = B_i = 0$.

Therefore, in linear perturbation theory, only the scalar perturbations of the metric are important for 
studying structure formation in the Universe. In our chosen Newtonian gauge, we end up with:
\beeq$ \label{eq:linearized-metric-scalar}
ds^{2}=-(1+2\Psi)dt^{2}+a^{2}(1-2\Phi)dx^{2}\,\,\,,
$
where the potentials $\Phi$ and $\Psi$ are functions of time and 
agree with the gauge invariant ``Bardeen`` potentials, since in this gauge: $ E = B = 0$ \mcite{Mukhanov Book}.
Let us now decompose the energy-momentum tensor $T_\munu $ also into a 
background ($\bar T$) and a perturbation part ($\delta T$):
\beeqalsp$\label{eq:perturbed-Tmunu}
T^0_0 &= -(\bar \rho + \delta \rho) \\
T^0_i &= (\bar \rho + \bar p)v_i \\
T^i_j &= (\bar p + \delta p)\delta^i_j + \Sigma_j^i \quad,
$
where $\Sigma_j^i \equiv T^i_j -\delta^i_j T^k_k/3$ is the traceless part of the energy momentum tensor and $\Sigma^i_i =0$.
Introducing \cref{eq:linearized-metric-scalar} and \cref{eq:perturbed-Tmunu}
into the Einstein equations \ref{eq:Einstein-field-equations}, we find to linear order:
\beeqal$
\nabla^2 \Phi -3\curH \left( \dot \Phi + \curH \Psi \right) &= 4 \pi G a^2 \delta\rho \label{eq:pertphipsi-1}\\
\nabla^2 \left(\dot \Phi + \curH \Psi \right) &= -4\pi G a^2 (\bar \rho + \bar p) \theta \label{eq:pertphipsi-2} \\
\begin{split}\label{eq:pertphipsi-3}
\ddot \Phi + \curH \left( \dot \Phi + 2 \dot \Psi \right) + \left( 2 \ddot{\curH} + \curH^2 \right) \Psi {} & \\
 \phantom{ \left( 2 \ddot{\curH} + \curH^2 \right) \Psi } + \frac{1}{3} \nabla^2 (\Psi - \Phi) & = 4\pi G a^2 \delta p  
\end{split}\\  
\nabla^2 \left( \Psi - \Phi \right) = 12 \pi G a^2 (\bar \rho + \bar p) \sigma \label{eq:pertphipsi-4}
$
with the peculiar velocity gradient $\theta \equiv \partial_i v^i $ and the anisotropic
stress perturbation $\sigma$ defined through $ (\bar p + \bar \rho)\nabla^2\sigma = -(\partial_i \partial^j - \frac{1}{3} \delta_i^j \nabla^2) \Sigma_j^i $.
Additionally to \crefrange{eq:pertphipsi-1}{eq:pertphipsi-4}, we still have the covariant conservation of the energy-momentum tensor: $\nabla_{\mu} T^\munu$,
which yields:
\beeqal$
\dot \delta &= -(1+w)(\theta - 3 \dot \Phi ) - 3 \curH (c^2_s - w) \delta \label{eq:deltadot-general}\\
\dot \theta &= -\curH (1-3w)\theta - \frac{\dot w}{1+w} \theta - \frac{c^2_s}{1+w} \nabla^2 \delta + \nabla^2 \sigma - \nabla^2 \Psi \label{eq:thetadot-general} \quad, 
$
where we have defined $c^2_s \equiv \delta p / \delta \rho $ and we now use the notation $\delta \equiv \delta \rho / \bar \rho$.
In order to close these set of equations, one needs to specify a sound speed and an equation of state and furthermore fix the anisotropic
stress perturbation $\sigma$. The anisotropic stress perturbation $\sigma$
is usually only sourced by relativistic particles, like photons 
and neutrinos in the early Universe.
For Cold Dark Matter (CDM), the equations simplify considerably by taking $\sigma = w = c_s^2 = 0$, as
we will see below in \cref{chap:nonlinear}.

If we go now to Fourier space, where $\nabla^2 \rightarrow -k^2$,
and $\partial_i v^i \rightarrow i k_i v_i = i k v $ (where the last equality is valid if the wavevector and
the velocity vector are aligned), then \cref{eq:pertphipsi-1} and \cref{eq:pertphipsi-2} can be combined
to give the relativistic Poisson equation:
\beeqc$\label{eq:relativistic-Poisson}
-k^2 \Phi = 4\pi G a^2 \bar{\rho} \Delta 
$
where we have defined the comoving relativistic perturbation $\Delta$ as:
\beeqp$\label{eq:relativistic-Delta}
 \Delta = \delta + \frac{3 \curH}{k} (1+w) i v
$
The Newtonian limit of these equations is obtained for non-relativistic particles, when $\sigma \ll v^2 \approx 0$ and we focus on the sub-horizon limit, in which the scales $k$ we are interested in, are much larger than the Hubble horizon, $k \gg \curH$.
In this limit we obtain from \cref{eq:pertphipsi-4} and \cref{eq:relativistic-Poisson}:
\beeqal$\label{eq:Newtonian-Poisson}
\Phi &= \Psi \\
\nabla^2 \Phi &= 4 \pi G a^2 \bar{\rho} \delta
$


%\begin{itemize}
%	\item Gravitational potentials $\Phi$ and $\Psi$.
%%	\item Very simple explanation of Gauge choice \comm{probably don't need it}
%	\item Newtonian Gauge
%%	\item Smallness of gravitational potentials \comm{what is this?}
%	\item Remark for later that in GR $\Psi=\Phi$
%	\item Weyl potential (lensing potential, null geodesics)
%	\item{The Newtonian limit}
%\end{itemize}



\input{Chapters/DE-MG}

%
%\begin{itemize}
%\item Differentiate between linear perturbations in different eras, just shortly
%\item Linear perturbations in matter dominated eras, newtonian gauge
%\item Fluid equations, full and linearized
%\end{itemize}


%----------------------------------------------------------------------------------------

%\section{Early Universe}
%\comm{We can probably delete this whole section}
%\subsection{Cosmic Microwave Background Radiation}
%\begin{itemize}
%\item Short introduction and importance of CMB
%\item Important constraints on parameters coming from CMB
%\item Constrain CDM alone
%\item Constrain initial power amplitude and tilt
%\item Constrain relativistic degrees of freedom
%\end{itemize}
%
%\subsection{Inflation} \comm{I would delete this}
%\begin{itemize}
%\item Inflation as a paradigm
%\item Flatness and horizon problems
%\item Inflation produces almost scale invariant spectrum
%\end{itemize}





